An investigation into using scagnostics to find interesting projections of high-dimensional data

### 2018-10-25 scagnostics.Rmd
Reference added.

  - smoothness: the index function should evolve smoothly between nearby projections, i.e. on an interpolated path we expect to observe slow and consistent change in the index function. \textcolor{red}{maybe can relate this to consistency consition developed by P. Hall, but don't understand this at the moment. In addition this relates to questions of optimiser choice and efficiency.} We evaluate smoothness both for transitions between noise variables and when moving from noise variables to an informative view in Section \ref{sec:smooth}.
    - distribution of index values: the distribution of index values when studying data points without structure should have a low variance, and be clearly separated from index values obtained when looking at structured data distributions. \textcolor{red}{I dont really understand this part. I think it can instead be related to the concept of squint angle.} We show examples in Section \ref{sec:dist}
  - ability to find fine and broad structure: \textcolor{red}{there are two different considerations, one  that relates to the concept of squint angle discussed in \cite{posse95b}. On the other hand should also consider detection of "fine" structure vs noise, this would be controlled e.g. by the allowed number of bins for the MINE index functions, additionally this will impact computing times, i.e. trade accuracy for efficiency}. We address these questions in Section \ref{sec:params} with an illustrative example.
  - rotation invariance: in the tour there is an assumed degeneracy between all projections obtained when rotating within the projected space only, for 2-d projections it means that we cannot assing x- and y- axis without ambiguity. A projection pursuit index should therefore be invariant under rotation of the basis in the projection plane. By costruction this is not the case for most index functions defined above. \textcolor{red}{One way around could be profiling the function over all possible rotations in the plane, howevere computing time may be a major concern in this case. Other examples where rotation invariance is important include image recognition and texture classification, rotation angle may be built in parameter for summary calculation, e.g. https://doi.org/10.1016/S0146-664X(75)80008-6, other example is interpolating over angles when building NN, see e.g. https://arxiv.org/pdf/1805.12301.pdf)}. We illustrate the impact of non invariance in Section \ref{sec:rot}. \textcolor{red}{consider frame interpolation rather than plane interpolation, which will avoid issues of rotation when interpolating.}
    - speed: the computation time is an important factor in particular when the guided tour is used to view the optimisation in real time. Note that longer computation times may be acceptable when first recording the optimisation and replaying the guided tour once the maximum has been identified. Computation times for some benchmark scenarios are compared in Section \ref{sec:timer}.

>>>>

  - smoothness: the index function should evolve smoothly between nearby projections, i.e. on an interpolated path we expect to observe slow and consistent change in the index function. \textcolor{red}{maybe can relate this to consistency consition developed by P. Hall, but don't understand this at the moment. In addition this relates to questions of optimiser choice and efficiency.} We evaluate smoothness both for transitions between noise variables and when moving from noise variables to an informative view in Section \ref{sec:smooth}.
    - distribution of index values: the distribution of index values when studying data points without structure should have a low variance, and be clearly separated from index values obtained when looking at structured data distributions. \textcolor{red}{I dont really understand this part. I think it can instead be related to the concept of squint angle \cite{barnett1981interpreting}.} We show examples in Section \ref{sec:dist}
  - ability to find fine and broad structure: \textcolor{red}{there are two different considerations, one  that relates to the concept of squint angle discussed in \cite{posse95b}. On the other hand should also consider detection of "fine" structure vs noise, this would be controlled e.g. by the allowed number of bins for the MINE index functions, additionally this will impact computing times, i.e. trade accuracy for efficiency}. We address these questions in Section \ref{sec:params} with an illustrative example.
  - rotation invariance: in the tour there is an assumed degeneracy between all projections obtained when rotating within the projected space only, for 2-d projections it means that we cannot assing x- and y- axis without ambiguity. A projection pursuit index should therefore be invariant under rotation of the basis in the projection plane. By costruction this is not the case for most index functions defined above. \textcolor{red}{One way around could be profiling the function over all possible rotations in the plane, howevere computing time may be a major concern in this case. Other examples where rotation invariance is important include image recognition and texture classification, rotation angle may be built in parameter for summary calculation, e.g. https://doi.org/10.1016/S0146-664X(75)80008-6, other example is interpolating over angles when building NN, see e.g. https://arxiv.org/pdf/1805.12301.pdf)}. We illustrate the impact of non invariance in Section \ref{sec:rot}. \textcolor{red}{consider frame interpolation rather than plane interpolation, which will avoid issues of rotation when interpolating.}
    - speed: the computation time is an important factor in particular when the guided tour is used to view the optimisation in real time. Note that longer computation times may be acceptable when first recording the optimisation and replaying the guided tour once the maximum has been identified. Computation times for some benchmark scenarios are compared in Section \ref{sec:timer}.
